### 2024-01-23 manuscript/02_methods.qmd

---
bibliography: references.bib
---

# Methods

## Oceanographic Sampling

Data were collected in collaboration with the Bermuda Atlantic Time-Series Study [@michaels1996; @lomas2013] on board the R/V Atlantic Explorer. Cruises were conducted at approximately monthly intervals. Rhizaria individuals were sampled using the Underwater Vision Profiler 5 [UVP5; @picheral2010], a tool which is well established to accurately quantify large Rhizaria [@biard2016; @stukel2018; @stukel2019; @biard2020; @barth2022; @drago2022; @llopismonferrer2022; @panaïotis2023]. The UVP5 was mounted to the sampling rosette and collected data autonomously on routine casts, from which only the downcast data are utilized. The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022. Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (Supplemental Figure 1; Supplemental Table 1). In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m). However, Rhizaria were only typically found in large abundances throughout the epipelagic and mesopelagic zones. As such, we limit this study to results from the upper 1000m of the water column.

A variety of biotic and abiotic data were collected during each BATS cruise. Briefly, we will explain the data utilized in this study. Salinity, temperature, Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were measured at high-frequencies on each cast, attached to the same CTD rosette as the UVP. On certain casts, niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths. On each cruise, flux estimates of total mass, carbon, and nitrogen were also collected using sediment traps (in the present study we utilize flux to the mesopelagic as the flux at 200m). Also primary productivity was estimated through measuring $C^{14}$ uptake rates from in-situ incubations. Full descriptions of the BATS sampling program, methods, and data are available freely online (https://bats.bios.asu.edu/bats-data/) and reviewed in [@lomas2013].

Environmental data were processed in a variety of ways to match the format of the Rhizaria abundance estimates (See below). CTD data were collected at higher frequency than the UVP, so these data were averaged within matching bins to the UVP5 data. Data from niskin bottles were first linearly interpolated with a 1m resolution for each cast where data were available. Then, because data were collected at different frequencies over the course of a cruise, the interpolated data were averaged across casts, then averaged into matching UVP5-sized bins. Primary productivity estimates were totaled within the euphotic zone to represent a "total euphotic productivity". Finally the data from the sediment trap deployed to 200m was used to represent "flux to the mesopelagic".

## Rhizaria imaging processing and quantification

Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa [cite]. Data were pre-sorted according to a . Taxonomic classification were done based on morphology exclusively. While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020]. Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: (https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf). Broadly, Rhizaria were classified as Foraminifera, Radiolarians (Acantharea or Collodaria), or a variety of Phaeodarian families (Figure 1). When identification could not be confidently made between a few candidate taxa, a less specific label was used. As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly phaeodaria but not distinguishable into a family.

The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$. However, identified rhizaria ranged from a 934$\mu m$ Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To confirm that the UVP5 was sampling adequately across all size ranges, an NBSS slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inrev] . However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded. The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals. In our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic [see @barth2022 for an extended discussion of UVP5 data processing). For the present study, Rhizarian abundances were estimated in 25m vertical bins, which offer a moderate sampling volume (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic) while still maintaining ecologically relevant widths. However, concentrations in a 25m bin would need to be greater than 2.428 indv. $m^{-3}$ and 3.912 indv. $m^{-3}$, in the epipealgic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inrev]. Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions. For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast. Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m). The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 indv. $m^{-2}$, 0.33 indv. $m^{-2}$, and 0.20 indv. $m^{-2}$ respectively.

## Modelling environmental controls of Rhizarian Abundance

Generalized Additive Models (GAMs) were used to assess the relationship between integrated rhizarian abudance and different environmental factors.