### 2024-01-25 manuscript/main.qmd

---
title: Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.
bibliography: references.bib
csl: ./deep-sea-research-part-i.csl
execute:
  echo: false
  warning: false
  message: false
author:
  - name: Alex Barth*
    affiliations:
    - name: University of South Carolina
      department: Biological Sciences
      address: 700 Sumter St 401, Columbia, SC 29208
      email: AB93@email.sc.edu
      phone: +1 (440) 668-8376
  - name: Leo Blanco-Berical
    affiliations:
    - name: Arizone State University
  - name: Rod Johnson
    affiliations: 
    - name: Bermuda Institute of Ocean Sciences
      address: Bermuda GE 01, UK
  - name: Joshua Stone
    affiliations:
    - name: University of South Carolina
      department: Biological Sciences
      address: 700 Sumter St 401, Columbia, SC 29208
format:
  pdf:
    template-partials:
      - "before-body.tex"
      - "_titlepage.tex"
    include-in-header:
      - "in-header.tex"
    number-sections: false
    linestretch: 2
    tbl-cap-location: top
    filters: [wordcount.lua]
    keep-tex: true
---

```{r}
#| child: ./01_intro.qmd
```

```{r}
#| child: ./02_methods.qmd
```

```{r}
#| child: ./03_results.qmd
```

```{r}
#| child: ./04_discussion.qmd
```

# References:
:::{#refs}
:::