### 2024-01-25 manuscript/main.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{hyphenat} % don't hyphenat titles
\usepackage{graphicx}
\usepackage{lineno}\linenumbers

% Add any tex header commands here
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leo Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Rhizaria in the oligotrophic exhibit clear temporal and vertical
variability.}
\author{Alex Barth* \and Leo Blanco-Berical \and Rod Johnson \and Joshua
Stone}
\date{}

\begin{document}
    \begin{titlepage}
% This is a combination of Pandoc templating and LaTeX
% Pandoc templating https://pandoc.org/MANUAL.html#templates
% See the README for help

\raggedleft % left align the title page
\rule{1pt}{\textheight} % Vertical line
\hspace{0.05\textwidth} % Whitespace between the vertical line and title page text
% Adjust num before \textwidth to move the block left or right
\begin{minipage}[b][\textheight][s]{0.85\textwidth}

\raggedright
% Title and subtitle
{\large\bfseries\nohyphens{Rhizaria in the oligotrophic exhibit clear
temporal and vertical variability.}}\\[2\baselineskip]

  % Authors	
% This hairy bit of code is just to get "and" between the last 2
% authors. See below if you don't need that
 {\large{Alex Barth*}}{\textsuperscript{1}}%
%
, 
 {\large{Leo Blanco-Berical}}{\textsuperscript{2}}%
%
, 
 {\large{Rod Johnson}}{\textsuperscript{3}}%
%
%
{ and \large{Joshua Stone}}%
{\textsuperscript{1}}%
%

% This is how to do it if you don't need the "and"

  %%%%%% Affiliations
\vspace{2\baselineskip}

\hangindent=1em
\hangafter=1
%
{1}.~{University of South Carolina}%
%
, %
{Biological Sciences}%
%
%
, %
{700 Sumter St 401, Columbia, SC 29208}%
%
\par\hangindent=1em\hangafter=1%
%
{2}.~{Arizone State University}%
%
%
\par\hangindent=1em\hangafter=1%
%
{3}.~{Bermuda Institute of Ocean Sciences}%
%
%
, %
{Bermuda GE 01, UK}%
%

  %%%%%% Correspondence
\vspace{1\baselineskip}

  %use \vfill instead to get the space to fill flexibly	
%\vspace{0.25\textheight} % Whitespace between the title block and the publisher

\vfill

%%%%%% Cover image

  % Whitespace between the title block and the tagline
\vspace{0.1\textheight}

\end{minipage}  \end{titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Rhizaria are an extremely diverse super-group of single celled organisms
consisting of several phyla including retaria (foraminifera and
radiolaria) and cercozoa. These organisms exist in a wide range of
habitats and are widely represented in plankton communities throughout
the global ocean. While the taxonomy of these organisms has recently
undergone several reclassifications (Biard, 2022a), their presence in
ocean ecosystems has been long known to oceanographers. Some of the
earliest records of their existence are from oceanographic expeditions
in the 19th century (Haekel, 1887). Rhizaria are unique members of the
plankton and protist community because they can reach large sizes (up to
several mm in diameter) and they construct intricate mineral skeletons
out of either silica, strontium, or calcium carbonate (Biard, 2022a;
Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
their noticeable morphology and global distribution, Rhizaria were
largely understudied throughout the 20th century. The bulk of modern
plankton research has focused hard-bodied crustacea which are
numerically dominant and easily sampled with nets and preservatives.
Fragile organisms like Rhizaria, were difficult to adequately study as
they can be destroyed through standard zooplankton sampling techniques.
A number of studies in the late 1900s did employ alternative techniques
to quantify Rhizaria including diaphragm pumps (Michaels, 1988) or
blue-water SCUBA collections (Bijma et al., 1990; Caron et al., 1995;
Caron and Be, 1984). However, the bulk of Rhizaria research was
constrained to sediment traps or paleontological studies of sediment
(Boltovskoy et al., 1993; Takahashi et al., 1983). Only recently has the
advent of molecular techniques and in-situ imaging tools ignited a
renewed focus on Rhizaria in pelagic ecosystems (Caron, 2016).

The wave of new data on Rhizaria has facilitated an improved
understanding of the significance in ocean ecosystem functions. Firstly,
taxonomists have been able to greatly refine the understanding of
evolutionary relationships amongst these diverse protists (Aurahs et
al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
revealed insights into the distributional patterns Llopis Monferrer et
al. (2022), ecological relationships (Decelle et al., 2012; Nakamura et
al., 2023), and contribution to biogeochemical fluxes (Guidi et al.,
2016; Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
approaches also have been used to quantify rhizarian contribution to
community metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizarian abundance
or biomass. In-situ imaging tools however, offer the ability to observe
organisms in the natural state and quantify their abundance (Barth and
Stone, in review; Ohman, 2019). While there were early applicaitons of
imaging tools to document Rhizaria (Dennett et al., 2002), Biard et al.
(2016)'s report from a global imaging dataset highlighted the importance
of Rhizaria to the total standing stock of marine carbon. Due to their
large sizes, ability to concentrate smaller particles (ballasting), and
the unique structure of their mineral skeletons, Rhizaria have the
potential to massively influence ocean biogeochemical cycling. A number
of studies have made large advances in estimating the contribution of
Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019;
Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica
(Biard et al., 2018; Llopis Monferrer et al., 2021), and strontium
(Decelle et al., 2013). Still, Rhizarian ecological roles are not well
understood (Biard, 2022a). This is a major challenge as it is critical
to understand the ecological role of plankton to fully incorporate them
into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit every different trophic
modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
2022a), yet their feeding modes can be quite varied. Phaeodarians
(family Cercozoa) are largely thought to be flux-feeders, collecting and
feeding on sinking particles (Nakamura and Suzuki, 2015; Stukel et al.,
2019). Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and contribute to total
primary productivity (Kimoto, 2015). Still, foraminifera are omnivorous,
possibly even predominately carnivorous with several studies suggesting
that they can be effective predators (Anderson and Bé, 1976; Gaskell et
al., 2019), majoritively consuming live copepods (Caron and Be, 1984).
Radiolaria have several lineages all which have some taxa who are well
known to host symbionts (Biard, 2022b). Amongst radiolarians, arguably
the most widespread are Collodaria who can be either large solitary
cells or form massive colonies, up to several meters in length (Swanberg
and Anderson, 1981). All known Collodaria species host dinoflagellate
symbionts (Biard, 2022b) and can contribute substantially to primary
productivity, particularly in oligotrophic ocean regions (Caron et al.,
1995; Dennett et al., 2002). This Collodaria-symbiont association has
been suggested as a reason for their high abundances throughout the
photic zone of oligotrophic environments globally (Biard et al., 2017,
2016). A few Acantharean (Radiolarian order) clades host algal symbionts
(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
exclusive relationship with Phaeocystis. However, globally Acantharea
are less abundant than Collodaria (Biard, 2022a) and contribute less to
total primary productivity (Michaels et al., 1995). This may be due to
the fact several clades of Acantharea are cyst-forming and strictly
heterotrophs (Biard, 2022b; Decelle et al., 2013). Furthermore, Mars
Brisbin et al. (2020) documented apparent predation behavior in
Acantharea near the surface, suggesting that there may be a large
reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizarian taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxa-specific studies of radiolarians suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022). Phaeodarians
alternatively, are generally found in the mesopelagic where
photosynthesis cannot occur but they can feed on sinking particles
(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
community, Biard and Ohman (2020) noted clear patterns in vertical
zonation which largely corresponded to different trophic roles. In the
oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
protist community, including Rhizaria partition along an autotroph and
mixotroph to heterotroph gradient with increasing depth in the water
column. Yet, few studies have made direct attempts to relate rhizaria
abundances to environmental factors (Biard and Ohman, 2020). In part,
this is due to the fact few studies have been able to sample Rhizaria in
the same location over a consistent timeframe (Boltovskoy et al., 1993;
Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
arguably the best method for quantifying rhizaria, consistently
throughout the full mesopelagic. Given this lack of information, there
are many unknowns with respect to Rhizarian ecology, seasonality and
phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
measured for multiple months (greater than 1 year) using an in-situ
imaging approach. With this dataset, we address two critical aims. 1)
Quantification of large Rhizaria throughout the epipelagic (0-200m) and
mesopelagic (200-1000m) over the course of an annual cycle. These data
were collected in the Sargasso Sea, and represents the first study of
its kind in an oligotrophic system. 2) We aim to test the hypothesis
that Rhizaria exhibit niche partitioning according to trophic roles.
This hypothesis makes several predictions, including vertical zonation,
as seen in prior studies, but also that environmental variables related
to trophic strategy will explain abundance patterns. Specifically,
autotrophic/mixotrophic taxa will correspond to variables related to
autotrophy (chl-a concentration, primary productivity, high \(O_2\) )
and other rhizaria will correspond to factors which promote heterotrophy
(particle concentration, flux, and low \(O_2\)).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{oceanographic-sampling}{%
\subsection{Oceanographic Sampling}\label{oceanographic-sampling}}

Data were collected in collaboration with the Bermuda Atlantic
Time-Series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
the R/V Atlantic Explorer. Cruises were conducted at approximately
monthly intervals. Rhizaria individuals were sampled using the
Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
which is well established to accurately quantify large Rhizaria (Barth
and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
rosette and collected data autonomously on routine casts, from which
only the downcast data are utilized. The UVP5 was deployed from
June-September 2019 then from October 2020 - January 2022. Casts were
filtered to only include data collected in the BATS region, far offshore
of Bermuda into the Sargasso Sea (Supplemental Figure 1). In general,
casts extended to either 200m, 500m, or 1200m deep, with a few extended
into the bathypelagic (4500m). However, Rhizaria were only typically
found in large abundances throughout the epipelagic and mesopelagic
zones. As such, we limit this study to results from the upper 1000m of
the water column.

A variety of biotic and abiotic data were collected during each BATS
cruise. Briefly, we will explain the data utilized in this study. The
UVP5 provided particle count data at a high-frequency from each cast.
Particle concentration was calculated from this data for all particles
184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
be reliably sampled by the UVP5's pixel resolution (\textgreater2px) and
the upper size range is representative of a potential prey field for
mesozooplankton (Whitmore and Ohman, 2021). Salinity, temperature,
Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were
measured at high-frequencies on each cast, attached to the same CTD
rosette as the UVP. On certain casts, niskin bottles were used to
collect bacterial abundance estimates (via epifluorescence microscopy)
as well as measure inorganic nutrients (\(NO_3\), and \(Si\) as
silicate/sicilic acid) at discrete depths. On each cruise, flux
estimates of total mass, carbon, and nitrogen were also collected using
sediment traps (in the present study we utilize flux to the mesopelagic
as the flux at 200m). Also primary productivity was estimated through
measuring \(C^{14}\) uptake rates from in-situ incubations. Full
descriptions of the BATS sampling program, methods, and data are
available freely online
\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}
and reviewed in (Lomas et al., 2013).

Environmental data were processed in a variety of ways to match the
format of the Rhizaria abundance estimates (See below). CTD data were
collected at higher frequency than the UVP, so these data were averaged
within matching bins to the UVP5 data. Data from niskin bottles were
first linearly interpolated with a 1m resolution for each cast where
data were available. Then, because data were collected at different
frequencies over the course of a cruise, the interpolated data were
averaged across casts, then averaged into matching UVP5-sized bins.
Primary productivity estimates were totaled within the euphotic zone to
represent a ``total euphotic productivity''. Finally the data from the
sediment trap deployed to 200m was used to represent ``flux to the
mesopelagic''.

\hypertarget{rhizaria-imaging-processing-and-quantification}{%
\subsection{Rhizaria imaging processing and
quantification}\label{rhizaria-imaging-processing-and-quantification}}

Individual vignettes of Rhizaria images were identified using the
classification platform Ecotaxa {[}cite{]}. Data were pre-sorted
according to a . Taxonomic classification were done based on morphology
exclusively. While there are sparse taxonomic guides for in-situ images
of rhizaria, identification largely relied on descriptions in (Nakamura
and Suzuki, 2015; Suzuki and Not, 2015; and Biard and Ohman, 2020).
Using the aforementioned sources and publicly available ecotaxa
projects, I constructed a guide which can be found here:
\url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
Broadly, Rhizaria were classified as Foraminifera, Radiolarians
(Acantharea or Collodaria), or a variety of Phaeodarian families (Figure
1). When identification could not be confidently made between a few
candidate taxa, a less specific label was used. As a result, we have
data from ``unidentified Rhizaria'', which typically were vignettes not
distinguishable between Aulacanthidae or Acantharea or ``unidentified
Phaeodaria'', which are clearly phaeodaria but not distinguishable into
a family.

\begin{figure}

{\centering \includegraphics{images/01_taxa.pdf}

}

\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
in lower right. All vignettes are same scale}

\end{figure}

The UVP5 samples at \textasciitilde15Hz rate as it descends the water
column and records the exact position of each particle larger than
600\(\mu m\). However, identified rhizaria ranged from a 934\(\mu m\)
Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To
confirm that the UVP5 was sampling adequately across all size ranges, an
NBSS slope was constructed to identify a drop-off which would indicate
poor-sampling at the small size range (Barth and Stone, in review;
Lombard et al., 2019) . However, it was evident from this analysis that
all size ranges were adequately sampled across the size range
(Supplemental Figure 2) so no data were excluded. The UVP5 reports the
exact depth at which a particle is recorded, however to estimate
abundance, observations must be binned over fixed depth intervals. In
our deployments had variable descent depths and speeds with more casts
descending to 500m than 1000m and descents quicker through the
epipelagic than the mesopelagic {[}see Barth and Stone (2022) for an
extended discussion of UVP5 data processing). For the present study,
Rhizarian abundances were estimated in 25m vertical bins, which offer a
moderate sampling volume (average 0.948\(m^3\) in the epipelagic and
0.589\(m^3\) in the mesopelagic) while still maintaining ecologically
relevant widths. However, concentrations in a 25m bin would need to be
greater than 2.428 indv. \(m^{-3}\) and 3.912 indv. \(m^{-3}\), in the
epipelagic and mesopelagic respectively, to fall below a 10\%
non-detection risk (Barth and Stone, in review; Benfield et al., 1996).
Because we typically observed many rhizaria taxa below these
concentrations, we present the 25m binned data to visualize broad-scale
average distributions. For quantifying and modelling Rhizaria
abundances, we present integrated abundance estimates, with each cast.
Due to the variable descent depths of the UVP, data are categorized as
epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
(500-1000m). The average sampling volume integrated through these
regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
non-detection thresholds at 0.30 indv. \(m^{-2}\), 0.33 indv.
\(m^{-2}\), and 0.20 indv. \(m^{-2}\) respectively. All UVP data
processing was done using the \texttt{EcotaxaTools} package in R (Barth
2023).

\hypertarget{modelling-environmental-controls-of-rhizarian-abundance}{%
\subsection{Modelling environmental controls of Rhizarian
Abundance}\label{modelling-environmental-controls-of-rhizarian-abundance}}

Generalized Additive Models (GAMs) were used to assess the relationship
between integrated Rhizarian abundance and different environmental
factors. GAMs offer the ability to model non-linear and non-monotonic
relationships, which can be particularly useful in assessing ecological
relationships (Wood, 2017) and have been successfully applied to
Rhizarian ecology (Biard and Ohman, 2020). The \texttt{mgcv} package
(Wood, 2001) was used to construct models relating environmental
parameters to each taxonomic group's integrated abundance estimates from
each cast. To select the most parsimonious model for each analysis, a
backwards step-wise approach was taken. First, a full model was fit
using any term which may be ecologically relevant. Terms were fit using
maximum likelihood with a double penalty approach on unnecessary smooths
(Marra and Wood, 2011). The smoothness parameter was restricted (k = 6)
to prevent overfitting the models. At each iteration of the backwards
step-wise procedure, the model term with the lowest F score (least
statistically significant) was removed. This was repeated until all
model terms were statistically significant or the \(R^2_{adj}\) was
substantially reduced. Models were fit for each region; epipelagic,
upper mesopelagic, and lower mesopelagic. In cases where observations
were too sparse for a given taxonomic grouping, models were not run.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{environmental-variability}{%
\subsubsection{Environmental
Variability}\label{environmental-variability}}

The BATS sampling region is southeast of Bermuda, situated in the North
Sargasso Sea and the North Atlantic oligotrophic ocean gyre. Due to the
sampling location, while the environmental conditions are generally low
in variation and oligotrophic, there is some seasonality and
considerable influence from mesoscale eddies (Lomas et al., 2013;
McGillicuddy et al., 1998). Variability in the water column structure
was visible during the study period (Figure 2). This is best evidenced
through the temperature profiles; In the late summer and early fall
there was a stratified water column with high temperatures in the
surface (\textless75m) (Figure 2A) and slightly elevated salinity
(Figure 2B). This warm, stratified period appeared more intense during
the few months sampled in 2019. In 2021, we observed the stratified
layer slowly dissipated into the winter months. There was a consistent
oxygen minimum zone (OMZ) located at about 750m deep (Figure 2C).
Although, February 2021 saw a notable downwelling event, likely due to a
passing cyclonic eddy. During this time, warmer, oxic water was plunged
deeper into the mesopelagic. This process was reversed in the spring
months (March, April) when mixing occurred throughout the epipelagic and
the surface was cooler and well mixed. Primary production was highest
during the spring mixing period, evidenced both by in-situ fluorescence
(Figure 2D) and productivity incubation experiments (Figure 3A).
Originating near the surface, the productivity peak moved deeper
throughout the spring and declined into the summer (Figure 2D). However,
there was a notable, yet smaller productivity bump in the late summer
and early fall (Figure 3A) which occurred deeper in the epipelagic
(Figure 2D). The particle concentration (184\(\mu m\) - 450\(\mu m\))
was closely coupled to chlorophyll-a patterns.

\begin{figure}

{\centering \includegraphics{images/02-environmental.pdf}

}

\caption{Environmental profiles across time-series of study period. Y
axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}

\end{figure}

Overall, particle concentration was high near the surface during the
2021 spring bloom, then sank throughout the water column attenuating
throughout the lower epipelagic (Figure 2E). Similarly, bacteria
abundance was closely linked to overall productivity, althrough there
was a more consistent layer near the top of the mesopelagic
(\textasciitilde250m) (Figure 2F). Concurrent with the secondary fall
production peak, there was also higher particle concentration and
bacterial abundance in the later summer and early fall. Interestingly,
while primary productivity estimates from July-August were not that
different between 2019 and 2021 (Figure 4A), chlorophyll-a florescence,
particle concentration, and bacterial abundance were much higher in
2019's summer/fall (Figure 2D-F). Inorganic nutrients (\(Si\) and
\(NO_3\)) were generally well stratified, with low concentrations in the
epipelagic and increasing throughout the mesopelagic. However, both
nutrients did vary vertically in accordance with the 2021 February
downwelling and spring mixing period (Figure 2G-H). Additionally, in the
late fall of 2021, \(Si\) concentrations were slightly elevated in the
mid-mesopelagic (Figure 2G).

Overall mass flux to the mesopelagic was highest during the 2021
February downwelling (Figure 3B). Generally, export was similarly high
during March, declining in April then increasing slightly throughout the
summer and early fall. While magnitude was slightly different, this
pattern was consistent with total mass, carbon and nitrogen fluxes
(Figure 3B-D). Higher mass, carbon and nitrogen flux also occurred in
the 2019 late summer - early fall period.

\begin{figure}

{\centering \includegraphics{images/03_prod-flux.pdf}

}

\caption{Primary productivity (A) and flux estimates from total mass
(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
month displayed in corresponding colors. Absent data are shown by grey
bar.}

\end{figure}

\hypertarget{rhizaria-abundance-and-distribution}{%
\subsubsection{Rhizaria abundance and
distribution}\label{rhizaria-abundance-and-distribution}}

Across all imaged mesozooplankton (\textgreater900\(\mu m\)), Rhizaria
comprised a considerable fraction of the total community. Considering
the total abundances of the observational period, Rhizaria comprised on
average, 42.6\% of all mesozooplankton abundance (Supplemental Figure
3). Copepods were the second most abundant, comprising 35.5\% and all
living mesozooplankton were 22\%. The large contribution of Rhizaria to
the mesozooplankton community is most prominent in the epipelagic, where
they accounted for 47\% of all mesozooplankton. In the mesopelagic
rhizaria were a smaller fraction, at 38\% in the upper layers (200-500m)
and 37\% in the deeper mesopelagic (500-1000m).

\begin{figure}

{\centering \includegraphics{images/04_vertical.pdf}

}

\caption{Average abundance of Rhizaria in 25m bins, across entire study
period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
(G), Coelodendridae (H).}

\end{figure}

Total average Rhizaria abundance had a bimodal distribution with respect
to depth. Total abundance was highest just below the surface (0-100m),
with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
Although, variation in depth binned abundance was large, likely due to
seasonal variability but also increased from the detection-risk
described in the methods. The vertical distribution pattern and
abundance varied considerably across taxonomic groups. Radiolarians were
some fo the most abundant taxa observed, particularly in the epipelagic
(Figure 4, Figure 5B). This pattern was led by Collodaria, whose
colonies were abundant in the upper epipelagic and declined into the top
of the mesopelagic (Figure 4C). Acantharea displayed a bimodal
distribution accounting for a large portion of the total Rhizaria
pattern (Figure 4B, Figure 5). Foraminifera had a similar bimodal
distribution, yet their overall average densities were much lower and
spread wider throughout the mesopelagic (FIgure 4E). Phaeodarian
families exhibited a wide range of vertical distribution patterns. The
most abundant, Aulacanthidae, also had a bimodal pattern but the density
was highest in the lower mesopelagic (Figure 4D). Aulosphaeridae had low
average density and was nearly homogenously distributed throughout the
water column, although slightly lower int eh epipelagic (Figure 4F).
Castanellidae were the only Phaeodarian who appeared to be effectively
restricted to the photic zone (Figure 4G). Alternatively, Coelodendridae
primarily occurred in the lower mesopelagic (Figure 4H). A few
individuals from the families Tuscaroridae and Medusettidae were also
observed in the mesopelagic, yet they were much rarer (data not shown).

Between the monthly cruises, Rhizaria integrated abundance varied in the
epipelagic. Highest average abundance occurred in June 2021 and was
lowest during the winter months (Figure 5A). The 2019 later summer -
fall period also had much higher integrated abundance than similar
months in 2021. While the majority of integrated abundance in the
epipelagic was consistently attributable to Collodaria, Acanthrean
abundance occurred sporadically and could account for a large portion of
the total in some months (Figure 5B). The mesopelagic integrated
abundance was much more consistent across monthly cruises, although
average abundance was notably higher in 2019 (Figure 5C-F). The
community composition in the mesopelagic was more diverse, mostly
comprised of Phaeodarians. However, Acantharea and unidentified Rhizaria
also were common members of the community (Figure 5D, 5F).

\begin{figure}

{\centering \includegraphics{images/05_seasonality.pdf}

}

\caption{Seasonality of Rhizarian integrated abundance for the
epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower
mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total
integrated abundance per monthly cruise colored by month. Right panels
(B, D, F) display community composition of each total abundance.}

\end{figure}

\hypertarget{environmental-drivers-of-rhizarian-abundance}{%
\subsubsection{Environmental Drivers of Rhizarian
Abundance}\label{environmental-drivers-of-rhizarian-abundance}}

For total Rhizarian integrated abundance the GAMs produced moderate fits
(\(R^2_{adj}\) = 0.406-0.603) (Table 1). In the epipelagic, there were
several significant predictor variables including inorganic nutrients
(\(NO_3\) and \(Si\)), water quality parameters (Salinity, DO), primary
production, and particle concentration (Table 1). However, the upper and
lower mesopelagic were exclusively explained by particle-related
variables (concentration and mass flux) (Table 1).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4286}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2857}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0952}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0833}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1071}}@{}}
\caption{Generalized Additive Model results for integrated total
Rhizarian abundance in different regions of the water
column.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p-value
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p-value
\end{minipage} \\
\midrule()
\endhead
\textbf{All Rhizaria Epipelagic} & Salinity & 2.0818 & 4.654 &
\textless0.001 \\
\(R^2_{adj}\)= 0.42 & DO & 0.8272 & 0.945 & 0.0136 \\
& Silica & 3.1206 & 15.75 & \textless0.001 \\
& NO3 & 2.9970 & 6.268 & \textless0.001 \\
& Primary Productivity & 1.8367 & 2.099 & \textless0.001 \\
& Particle Concentration & 0.9282 & 2.566 & \textless0.001 \\
& & & & \\
\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6847 & 0.431 &
0.073 \\
\(R^2_{adj}\)= 0.406 & Avg Mass Flux & 1.5426 & 0.923 & 0.045 \\
& Particle Concentration & 3.2957 & 20.89 & \textless0.001 \\
& & & & \\
\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 1.6632 & 1.824
& 0.002 \\
\(R^2_{adj}\)= 0.603 & Particle Concentration & 0.7694 & 0.662 &
0.027 \\
\bottomrule()
\end{longtable}

GAMs for individual taxa were much less consistent in their fits (Table
2). This is likely in part due to the high number of non-observations
for certain taxa. Note that due to low abundances, GAMS were not
constructed for Tuscaroridae or Medusettidae. Furthermore no significant
terms were found for a model with Aulosphaeridae in the epipelagic nor
Foraminifera in the mesopelagic.

\begin{figure}

{\centering \includegraphics{images/06_epi-partials.pdf}

}

\caption{Partial effects of smooth terms in taxa-specific GAM models
from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
(E).}

\end{figure}

Epipelagic Acantharea were explained by several predictor vairables and
had a good fit (\(R^2_{adj}\) = 0.53, Table 2). Most notable smooths
however, were mass flux and particle concentration, which had a weak
positive association (Figure 6A), with July 2021 as a clear outlier
where Acantharean abundances were high in the epipelagic despite lower
fluxes and particle concentrations (Figure 5). Foraminifera had a good
fitting GAM in the epipelagic (\(R^2_{adj}\) = 0.445). There were
several significant explanatory variables, althoguht the clearest
pattern was observed of high temperatures associated with more
Foraminifera abundance (Table 2, Figure 6B). Epipelagic Aulacanthidae
similarly had several predictor variables which were significant,
including both water quality parameters and particle/flux predictors
(Table 2). Interestingly, Aulacanthidae had primary production as a
significant predictor, yet there was not a clear association (Figure
6C). There was a fit for Collodaria in the epipelagic (\(R^2_{adj}\) =
0.16), although there was a logit-like relationship where higher
abundances tended to occur during higher DO conditions in the surface
waters (Figure 6D). Castanellidae also had similarly poor fits in the
epipelagic (\(R^2_{adj}\) = 0.124) (Table 2, Figure 6E).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4471}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2824}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1059}}@{}}
\caption{Taxa-specific generalized additive models for different regions
of the water column.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p-value
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p-value
\end{minipage} \\
\midrule()
\endhead
\textbf{Acantharea Epipelagic} & Salinity & 2.264 & 4.113 &
\textless0.001 \\
R2adj=0.53 & O2 & 2.579 & 6.712 & \textless0.001 \\
& Avg Mass Flux & 4.810 & 22.81 & \textless0.001 \\
& Bacteria \#/L & 1.599 & 1.317 & 0.0117 \\
& Particle Concentration & 0.853 & 1.158 & 0.0087 \\
& & & & \\
\textbf{Acantharea Upper Mesopelagic} & Avg C Flux & 1.296 & 0.621 &
0.0439 \\
R2adj=0.231 & Avg N Flux & 1.619 & 1.283 & 0.0026 \\
& Particle Concentration & 0.952 & 3.886 & \textless0.001 \\
& & & & \\
\textbf{Acantharea Lower Mesopelagic} & Temperature & 2.076 & 4.155 &
\textless0.001 \\
R2adj=0.509 & Avg N Flux & 1.494 & 1.216 & 0.0113 \\
& Primary Productivity & 0.766 & 0.648 & 0.0253 \\
& Particle Concentration & 2.037 & 10.03 & \textless0.001 \\
\textbf{Aulacanthidae Epipelagic} & Salinity & 0.792 & 0.757 & 0.0241 \\
R2adj=0.251 & Avg N Flux & 0.869 & 1.324 & 0.0018 \\
& Primary Productivity & 2.312 & 7.704 & \textless0.001 \\
& Particle Concentration & 2.008 & 2.200 & 0.0017 \\
& & & & \\
\textbf{Aulacanthidae Upper Mesopelagic} & Particle Concentration &
2.832 & 9.802 & \textless0.001 \\
R2adj = 0.158 & & & & \\
& & & & \\
\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.622 & 1.991
& 0.002 \\
R2adj=0.298 & Particle Concentration & 2.123 & 6.706 & \textless0.001 \\
& & & & \\
\textbf{Aulosphaeridae Upper Mesopelagic} & Particle Concentration &
2.653 & 13.06 & \textless0.001 \\
R2adj=0.2 & & & & \\
& & & & \\
\textbf{Aulosphaeridae Lower Mesopelagic} & Particle Concentration &
0.972 & 6.248 & \textless0.001 \\
R2adj=.147 & & & & \\
& & & & \\
\textbf{Castanellidae Epipelagic} & Avg C Flux & 1.462 & 0.816 &
0.0421 \\
R2adj = 0.124 & Primary Productivity & 0.771 & 0.665 & 0.0247 \\
& Particle Concentration & 3.956 & 4.623 & \textless0.001 \\
& & & & \\
\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.822 & 0.919 &
0.0183 \\
R2adj=.113 & Particle Concentration & 0.970 & 6.208 & \textless0.001 \\
& & & & \\
\textbf{Coelodendridae Lower Mesopelagic} & Particle Concentration &
1.773 & 4.873 & \textless0.001 \\
R2adj=.133 & & & & \\
& & & & \\
\textbf{Collodaria Epipelagic} & Salinity & 0.925 & 2.267 &
\textless0.001 \\
R2adj=0.16 & O2 & 2.015 & 2.217 & 0.002 \\
& Bacteria \#/L & 1.843 & 2.100 & 0.002 \\
& & & & \\
\textbf{Foraminifera Epipelagic} & Temperature & 4.414 & 8.789 &
\textless0.001 \\
R2adj=0.445 & O2 & 1.603 & 1.287 & 0.0111 \\
& Avg N Flux & 2.512 & 3.396 & \textless0.001 \\
& Primary Productivity & 0.824 & 0.920 & 0.0153 \\
& Particle Concentration & 0.919 & 2.257 & \textless0.001 \\
\bottomrule()
\end{longtable}

\begin{figure}

{\centering \includegraphics{images/07_upmeso-partial.pdf}

}

\caption{Partial effects of smooth terms in taxa-specific GAM models
from the upper mesopelagic (200-500m). Effects are grouped by taxa;
Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
(D).}

\end{figure}

In the upper mesopelagic (200-500m), abundances were generally low
(Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae,
Aulosphaeridae, and Coelodendridae (Table 2). All these models had
generally poor fits (\(R^2_{adj}\) \textless{} 0.25). Yet, for all upper
mesopelagic models, particle concentration was a significant explanatory
variable (Table 2, Figure 7). Carbon flux was significant for Acantharea
and nitrogen flux was significant for both Acantharea and Coelodendridae
(Table 2, Figure 7A,D). The lower mesopelagic also had generally poor
GAM fits for taxa specific models (\(R^2_{adj}\) \textless0.3), with the
exception of Acantharea (\(R^2_{adj}\) = 0.509). Acantharea in the lower
mesopelagic was most clearly positively associated with particle
concentration and nitrogen flux, as well as temperature to a slight
degree (Figure 8A). For all Phaeodarians with a significant model,
particle concentration was a main predictor variable (Table 2, Figure
8B-D). Aulacanthidae had the best fitting model of the Phaeodarians
(\(R^2_{adj}\) = 0.298), which also included mass flux as a
statistically significant smooth (Figure 8B).

\begin{figure}

{\centering \includegraphics{images/08_lomeso-partials.pdf}

}

\caption{Partial effects of smooth terms in taxa-specific GAM models
from the lower mesopelagic (500-1000m). Effects are grouped by taxa;
Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
(D).}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{overall-rhizarian-abundance-and-patterns}{%
\subsubsection{Overall Rhizarian abundance and
patterns}\label{overall-rhizarian-abundance-and-patterns}}

In the epipelagic Rhizaria exhibited a notable seasonal pattern.
Rhizarian abundances were higher in the summer months and lower during
the winter. During a prior time period, Blanco-Bercial et al. (2022)
noted that there is considerable seasonality in the community
composition of all protists. Despite the seasonality of total Rhizarian
abundance, community composition was relatively consistent, with
Collodaria representing the bulk of the community. It should be noted
that the overall taxonomic resolution of the UVP5 is fairly low, so
there may be a switching of species within the broad groups identified
in this study which were not captured. Throughout the mesopelagic,
month-to-month variation in 2021 was relatively low. Again this is
consistent with observations from metabarcoding of the whole protist
community in the same study region (Blanco-Bercial et al., 2022). This
finding is not surprising as the overall seasonal variation in
environmental conditions in this region were low.

Overall Rhizaria were the most commonly identified group of
mesozooplankton throughout the study period. We do note that the UVP5
commonly captures Trichodesmium colonies, yet these were excluded in
this comparison as they are strictly autotrophs. It should be noted that
previous work has suggested that avoidance behavior with the UVP is
possible, at times likely, for visual and highly mobile zooplankton
(Barth and Stone, 2022). Thus, the percent contribution reported here
(42.7\%) of Rhizaria to the total mesozooplankton community may be
inflated due to under sampling of organisms such as Euphausiids and
Chaetognaths which has quick escape responses. Regardless, it is worth
noting that in the same region, with data collected in 2012 and 2013
using similar calculation methods, Biard et al. (2016) estimated
Rhizaria only contribute 15\% of the total mesozooplankton community in
the upper 500m. Likely, Rhizaria display considerable interannual
variability. In the present study, we noticed considerably higher
Rhizarian abundance throughout the water column in 2019 compared to
2021. While this may have been driven by increased mass flux (Figure 3),
more information is needed to truly understand the magnitude by which
Rhizaria can vary interannually.

\hypertarget{vertical-structure-and-trophic-roles}{%
\subsubsection{Vertical Structure and Trophic
Roles}\label{vertical-structure-and-trophic-roles}}

In this study we present a clear pattern of vertical zonation between
different Rhizaria groups. Largely, the taxonomic composition and
vertical positioning were similar to Rhizaria zonation in the California
Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
that the secondary abundance peak reported in the present study is
lower. This is likely due to the more oligotrophic nature of the study
region, were the euphotic zone penetrates deeper into the water column.
Most prevalent in the epipelagic were Collodaria. These mixotrophic
Radiolarians have long been reported to contribute to primary
productivity in the euphotic zone (Dennett et al., 2002; Michaels et
al., 1995). Collodaria are thought to be particularly successful
globally in oligotrophic regions due to their photosymbiotic
relationships (Biard et al., 2017, 2016). We observed the highest
abundance of Collodaria during June 2021, supporting the notion they can
thrive during the typically low-nutrient conditions of summer
stratification. However, Collodaria also increased during the spring
mixing period, suggesting that they can thrive during conditions which
may typically be thought to favor autotrophs. Furthermore, while
Collodaria were primarily absent from below 250m, there were a few
instances of deeper colonies being observed. Global investigations of
polycistine flux, suggest that deep-Collodaria in Oligotrophic regions
may be a consequence of isothermal submersion (Boltovskoy, 2017).
Another effectively exclusively epipelagic Rhizaria was the Phaeodarian
family of Castanellidae. All Phaeodarians are thought to be fully
heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
studies, including this one, report Castanellidae to be typically found
in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
Zasko and Rusanov, 2005). It should be considered that perhaps
Castanellidae specializes in feeding sinking particles directly at the
base of the epipelagic. Given it's smaller size (Nakamura and Suzuki,
2015), Castanellidae does not need a large diameter to efficiently flux
feed at the typically particle rich region of the lower epipelagic. Both
Castanelldiae and Collodaria had poor fitting GAMs. This is somewhat of
a surprise for Collodaria who had large abundance. However, given the
consistency of their abundance, it may be that this study did not
capture a wide enough range of conditions for describing Collodaria's
preferred niche.

The mesopelagic generally was home to known heterotrophic organisms,
particularly for those which were constrained to exclusively occupy
deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
observation of an auto-/mixotroph to heterotroph gradient in the protist
community. The upper mesopelagic interestingly had relatively low total
abundance. This low-abundance region likely reflects the dynamics of
productivity and export throughout the water column. While productivity
and thus sinking particles for flux feeders are high in the euphotic
zone, much of this is attenuated throughout the epipelagic. So, while
the base of the epipelagic may provide a rich feeding environment for
Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F),
the region from 200-500m might be otherwise food poor. Perhaps it is
more advantageous for Rhizaria to situate deeper, in darker regions of
the twilight zone. Also it should be noted that Phaeodarians utilize
silica to build their opaline tests, and silica concentrations started
to increase around 500m (Figure 2G). Although \(Si\) was not a
significant smooth for any taxa-specific model, this lack of association
might be due to the overall lack of variation of \(Si\) between
integrated abundance of each cast. Aulosphaeridae was only found to have
significant relationships, although weak fits, to particle concentration
in the mesopelagic. In our study, while consistently observed, overall
abundances of Aulosphaeridae were very low. In the Pacific Ocean, on
California's Coast, much higher abundances of Aulosphaeridae have been
reported (Biard and Ohman, 2020; Zasko and Rusanov, 2005) and they have
massive potential to impact silica export (Biard et al., 2018).
Coelodendridae were also seemingly restricted to the deeper section of
the mesopelagic. This is interesting given that in the California
Current, (Biard and Ohman, 2020) found a bimodal distribution in
Coelodendridae. There are several morphotypes corresponding to different
taxa of Coelodendridae (Biard and Ohman, 2020; Nakamura and Suzuki,
2015). So it may be that only a few types of Coelodendridae were
observed in this study, while the epipelagic variety was not.
Alternatively, the lower epipelagic of the California Current may
provide adequate habitat for Coelodendridae, which is not available in
the oligotrophic Sargasso Sea.

A number of taxa were found to have a bimodal distribution, with
considerable populations in both the epipelagic and mesopelagic.
Aulacanthidae had a bimodal distribution, although abundances were
highest in the lower mesopelagic. Foraminifera also had a bimodal
distribution. Some lineages of Foraminifera are known to host
photosymbionts (Biard, 2022a; Kimoto, 2015), however they are also
efficient predators commonly seen throughout the mesopelagic (Caron and
Be, 1984; Gaskell et al., 2019). Thus it is not surprising to find their
presence in both locations of the water column. Foraminifera are also
known to vary their vertical distribution across their life cycle in
phase with lunar cycles (Biard, 2022a; Bijma et al., 1990; Gaskell et
al., 2019; Kimoto, 2015). However, the sampling scheme of the BATS
program does not capture this frequency and was not investigated in the
present study.

Acantharea also had a bimodal distribution, with much larger abundances
than Aulacanthidae or Foraminifera. Most prior studies of Acantharea
vertical distribtion found them concentrated in near surface layers of
the water column Zasko and Rusanov (2005). This would support the
paradigm that large Acantharean abundances may be supported by their
mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
While the UVP5 images cannot distinguish between mixotrophic and
heterotrophic Acantharea, the GAMs constructed for Acantharean abundance
found positive associations with particle concentration and mass flux,
suggesting a higher reliance on heterotrophy. Recently Mars Brisbin et
al. (2020) described apparent predator behavior amongst near-surface
Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
heterotrophic. Yet, it should be noted in the Sargasso Sea, both
heterotrophic and symbiotic lineages of Acantharea have been reported
(Blanco-Bercial et al., 2022). Additionally, Michaels (1988) noted that
the majority of Acantharea (by abundance) were smaller than
160\(\mu m\). While that estimate may be inflated due to inability to
capture larger cells, small Acantharea were not captured in the present
study. Thus, trophic strategy may shift based on sizes of Acantharea.

Decelle et al. (2013) proposed a hypothetical life cycle for
cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life
cycle suggests that epipelagic Acantharea are adult populations, which
form cysts that sink into the mesopelagic, then reproduce and raise.
Furthermore, given that horizontal transfer of symbionts between
generations of Acantharea is unlikely, the newly spawned mesopelagic
Acantharea are not necessarily required to rapidly return to the photic
zone. This hypothesis predicts that Acanthareans in the mesopelagic
would be smaller (Decelle et al., 2013). Mars Brisbin et al. (2020)
provided some support for this hypothesis, with a significant decrease
in Acantharea sizes with depth. Although, the authors also observed low
abundances in the mesopelagic and noted that the smaller sizes may be
due to less food availability (Mars Brisbin et al., 2020). Since food is
more scarce in the mesopelagic there is is less nutritional quality, yet
flux feeders would likely grow larger to increase their feeding range
(Biard and Ohman, 2020). In the data collecting in this study,
Acanthareans in the mesopelagic were significantly smaller than the
epipelagic, despite the other bimodal taxa (Foraminifera and
Aulacanthidae) being significantly larger with depth (Figure 9, Wilcoxon
rank sum text p-value \textless{} 0.001). This provides added support
for the hypothesis that cyst-forming Acantharea may utilize different
sections of the water column throughout their life cycle. However to
further investigate this, more work is needed with higher temporal and
taxonomic resolution.

\begin{figure}

{\centering \includegraphics{images/09_size-comp.pdf}

}

\caption{Comparison of average sizes (ESD) amongst Rhizarian taxa which
occurred throughout the water column.}

\end{figure}

\hypertarget{conclusions-and-considerations}{%
\subsubsection{Conclusions and
Considerations}\label{conclusions-and-considerations}}

This study provides a detailed look at Rhizarian abundance over time
throughout the water column in a major oligotrophic gyre. We show that
their abundances are generally related to particle concentration and
flux, although lack of environmental variability may have reduced the
fit of our GAMs. Considering the potential role of Rhizaria in the
biological carbon pump, they may have a somewhat mixed role. In the
shallower regions, where smaller Rhizaria are abundant, they may be an
attenuating force on sinking particles (Stukel et al., 2019). It should
be noted that in our study, we focused on the ``small'' particle
concentration field (\textless450\(\mu m\)), and these particles are
generally slower sinking than large particles. However, once consumed
and repackaged by larger Rhizaria, they can sink quicker and contribute
more to overall flux (Michaels, 1988). Thus, Rhizaria may act as a
ballasting mechanism. However, this is largely speculation, to truly
test this, more work is needed measuring Rhizarian flux.

The vertical partitioning documented in this study do support the
hypothesis that mixotrophic rhizaria will occupy shallower waters while
deeper waters are dominated by heterotrophy. However the degree to which
mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus
symbiosis is uncertain. Collodaria were recorded as consistent and
dominant members of the near surface region. These organisms have the
potential to contribute considerably to the otherwise low productivity
of oligotrophic regions. However, their role in food webs is not well
understood. While this study represents a step-forward in our
understanding of Rhizarian ecology, there are several known aspects of
Rhizarian and protist ecology which are not captured in this study.
Continued research on Rhizaria is much needed to better understand their
ecology. Particularly extended descriptive work to capture interannual
patterns. Also work defining biotic interactions, feeding rates,
productivity, and life history are all rich fields of interest in
Rhizaria.

\hypertarget{references}{%
\section{References:}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
Marine Protist Symbioses. Acta Protozoologica 53, 29--38.
\url{https://doi.org/10.4467/16890027AP.13.0019.1116}

\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA PELAGICA}}
(d'ORBIGNY). The Biological Bulletin 151, 437--449.
\url{https://doi.org/10.2307/1540498}

\leavevmode\vadjust pre{\hypertarget{ref-aurahs2009}{}}%
Aurahs, R., Grimm, G.W., Hemleben, V., Hemleben, C., Kucera, M., 2009.
Geographical distribution of cryptic genetic types in the planktonic
foraminifer Globigerinoides ruber. Molecular Ecology 18, 1692--1706.
\url{https://doi.org/10.1111/j.1365-294X.2009.04136.x}

\leavevmode\vadjust pre{\hypertarget{ref-barth-inrev}{}}%
Barth, A., Stone, J., in review. Understanding the picture: The promis
and challenges of in-situ imagery data in the study of plankton ecology.

\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
Barth, A., Stone, J., 2022.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.898057}{Comparison
of an in situ imaging device and net-based method to study
mesozooplankton communities in an oligotrophic system}. Frontiers in
Marine Science 9.

\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
R., Copley, N.J., 1996. Video Plankton Recorder estimates of copepod,
pteropod and larvacean distributions from a stratified region of Georges
Bank with comparative measurements from a MOCNESS sampler. Deep Sea
Research Part II: Topical Studies in Oceanography 43, 1925--1945.
\url{https://doi.org/10.1016/S0967-0645(96)00044-6}

\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
\url{https://doi.org/10.1002/9780470015902.a0029469}

\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
Environmental Microbiology 24, 2179--2200.
\url{https://doi.org/10.1111/1462-2920.16004}

\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
Pesant, S., Stemmann, L., Not, F., 2017. Biogeography and diversity of
Collodaria (Radiolaria) in the global ocean. The ISME Journal 11,
1331--1344. \url{https://doi.org/10.1038/ismej.2017.12}

\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
Significance of Giant Phaeodarians (Rhizaria) to Biogenic Silica Export
in the California Current Ecosystem. Global Biogeochemical Cycles 32,
987--1004. \url{https://doi.org/10.1029/2018GB005877}

\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
protists (Rhizaria) into the twilight zone revealed by in situ imaging.
Limnology and Oceanography 65, 2583--2602.
\url{https://doi.org/10.1002/lno.11472}

\leavevmode\vadjust pre{\hypertarget{ref-biard2015}{}}%
Biard, T., Pillet, L., Decelle, J., Poirier, C., Suzuki, N., Not, F.,
2015. Towards an integrative morpho-molecular classification of the
collodaria (polycystinea, radiolaria). Protist 166, 374--388.
\url{https://doi.org/10.1016/j.protis.2015.05.002}

\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
H., Gorsky, G., Guidi, L., Kiko, R., Not, F., 2016. In situ imaging
reveals the biomass of giant protists in the global ocean. Nature 532,
504--507. \url{https://doi.org/10.1038/nature17652}

\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
Bijma, J., Erez, J., Hemleben, C., 1990.
\href{https://epic.awi.de/id/eprint/6096/1/Bij1990a.pdf}{Lunar and
semi-luan reproductive cycles in some spinose planktonic foraminifers}.
Journal of Foraminiferal Research 20, 117--127.

\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
S.J., Curry, R., 2022.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.897140}{The
protist community traces seasonality and mesoscale hydrographic features
in the oligotrophic sargasso sea}. Frontiers in Marine Science 9.

\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
polycystina (protista) in the world ocean: Living ranges, isothermal
submersion and settling shells. Journal of Plankton Research 39,
330--349. \url{https://doi.org/10.1093/plankt/fbx003}

\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
radiolaria and other shelled plankters in the eastern equatorial
atlantic at 853 m: seasonal variations and polycystine species-specific
responses. Deep Sea Research Part I: Oceanographic Research Papers 40,
1863--1895. \url{https://doi.org/10.1016/0967-0637(93)90036-3}

\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
Caron, D.A., 2016.
\href{https://www.nature.com/articles/nature17892}{The rise of rhizaria
\textbar{} nature}. Nature 532, 444--445.

\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
THE SPINOSE PLANKTONIC FORAMINIFER. BULLETIN OF MARINE SCIENCE 35.

\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
productivity by symbiont-bearing planktonic sarcodines (acantharia,
radiolaria, foraminifera) in surface waters near bermuda. Journal of
Plankton Research 17, 103--129.
\url{https://doi.org/10.1093/plankt/17.1.103}

\leavevmode\vadjust pre{\hypertarget{ref-cavalier-smith2018}{}}%
Cavalier-Smith, T., Chao, E.E., Lewis, R., 2018. Multigene phylogeny and
cell evolution of chromist infrakingdom Rhizaria: contrasting cell
organisation~of sister~phyla Cercozoa and Retaria. Protoplasma 255,
1517--1574. \url{https://doi.org/10.1007/s00709-018-1241-1}

\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
McIlvin, M.R., Lopez, P.Z., Barth, A., Stone, J., Alanis, B.A., Chan,
E.W., Breier, J.A., Jakuba, M.V., Johnson, R., Alexander, H., Saito,
M.A., 2023. Microeukaryote metabolism across the western North Atlantic
Ocean revealed through autonomous underwater profiling.
\url{https://doi.org/10.1101/2023.11.20.567900}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
Planktonic Protists, in: Ohtsuka, S., Suzaki, T., Horiguchi, T., Suzuki,
N., Not, F. (Eds.),. Springer Japan, Tokyo, pp. 465--500.
\url{https://doi.org/10.1007/978-4-431-55130-0_19}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
F., De Vargas, C., Lampitt, R., Not, F., 2013. Diversity, Ecology and
Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the Oceans.
PLoS ONE 8, e53598. \url{https://doi.org/10.1371/journal.pone.0053598}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
Multiple microalgal partners in symbiosis with the acantharian
Acanthochiasma sp. (Radiolaria). Symbiosis 58, 233--244.
\url{https://doi.org/10.1007/s13199-012-0195-x}

\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
2002. Video plankton recorder reveals high abundances of colonial
radiolaria in surface waters of the central north pacific. Journal of
Plankton Research 24, 797--805.
\url{https://doi.org/10.1093/plankt/24.8.797}

\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
Carlotti, F., Coppola, L., Guidi, L., Hauss, H., Karp-Boss, L., Lombard,
F., Mcdonnell, A.M.P., Picheral, M., Rogge, A., Waite, A.M., Stemmann,
L., Kiko, R., 2022. Global Distribution of Zooplankton Biomass Estimated
by In Situ Imaging and Machine Learning. Frontiers in Marine Science 9.
\url{https://doi.org/10.3389/fmars.2022.894372}

\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
measurements of planktonic foraminifera in the california current
system. Journal of Foraminiferal Research 49, 390--404.
\url{https://doi.org/10.2113/gsjfr.49.4.390}

\leavevmode\vadjust pre{\hypertarget{ref-guidi2016}{}}%
Guidi, L., Chaffron, S., Bittner, L., Eveillard, D., Larhlimi, A., Roux,
S., Darzi, Y., Audic, S., Berline, L., Brum, J.R., Coelho, L.P.,
Espinoza, J.C.I., Malviya, S., Sunagawa, S., Dimier, C., Kandels-Lewis,
S., Picheral, M., Poulain, J., Searson, S., Stemmann, L., Not, F.,
Hingamp, P., Speich, S., Follows, M., Karp-Boss, L., Boss, E., Ogata,
H., Pesant, S., Weissenbach, J., Wincker, P., Acinas, S.G., Bork, P.,
Vargas, C. de, Iudicone, D., Sullivan, M.B., Raes, J., Karsenti, E.,
Bowler, C., Gorsky, G., 2016. Plankton networks driving carbon export in
the oligotrophic ocean. Nature 532, 465--470.
\url{https://doi.org/10.1038/nature16942}

\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
Scharek, R., Vaulot, D., Landry, M.R., Not, F., 2019. High contribution
of Rhizaria (Radiolaria) to vertical export in the California Current
Ecosystem revealed by DNA metabarcoding. The ISME Journal 13, 964--976.
\url{https://doi.org/10.1038/s41396-018-0322-7}

\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
Not, F., Fernández, D., Gourvil, P., Bilewitch, J., Hulston, D.,
Pinkerton, M., Nodder, S.D., 2022. Planktonic protist diversity across
contrasting subtropical and subantarctic waters of the southwest
pacific. Progress in Oceanography 206, 102809.
\url{https://doi.org/10.1016/j.pocean.2022.102809}

\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
Haekel, E., 1887.
\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
on the radiolaria collected by h.m.s. Challenger during the years
1873-1876, plates, by ernst haeckel}.

\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
and depth distribution of a mesopelagic foraminifer, Hastigerinella
digitata, in Monterey Bay, California. Limnology and Oceanography 56,
562--576. \url{https://doi.org/10.4319/lo.2011.56.2.0562}

\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
K., Harada, N., Fujiki, T., 2019. Phaeodaria: An Important Carrier of
Particulate Organic Carbon in the Mesopelagic Twilight Zone of the North
Pacific Ocean. Global Biogeochemical Cycles 33, 1146--1160.
\url{https://doi.org/10.1029/2019GB006258}

\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
\url{https://doi.org/10.1007/978-4-431-55130-0_7}

\leavevmode\vadjust pre{\hypertarget{ref-lampitt2009}{}}%
Lampitt, R.S., Salter, I., Johns, D., 2009. Radiolaria: Major exporters
of organic carbon to the deep ocean. Global Biogeochemical Cycles 23.
\url{https://doi.org/10.1029/2008GB003221}

\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
Lee, J.J., 2006.
\href{https://dalspace.library.dal.ca/bitstream/handle/10222/78255/VOLUME\%2042-NUMBER\%202-2006-PAGE\%2063.pdf?sequence=1}{Algal
symbiosis in larger foraminifera}. Symbiosis 42, 63--75.

\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
M., Elineau, A., Guidi, L., Leynaert, A., Tréguer, P.J., Not, F., 2022.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.895995}{Siliceous
rhizaria abundances and diversity in the mediterranean sea assessed by
combined imaging and metabarcoding approaches}. Frontiers in Marine
Science 9.

\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2021}{}}%
Llopis Monferrer, N., Leynaert, A., Tréguer, P., Gutiérrez-Rodríguez,
A., Moriceau, B., Gallinari, M., Latasa, M., L'Helguen, S., Maguer,
J.-F., Safi, K., Pinkerton, M.H., Not, F., 2021. Role of small Rhizaria
and diatoms in the pelagic silica production of the Southern Ocean.
Limnology and Oceanography 66, 2187--2202.
\url{https://doi.org/10.1002/lno.11743}

\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
Carlson, C.A., 2013. Two decades and counting: 24-years of sustained
open ocean biogeochemical measurements in the Sargasso Sea. Deep Sea
Research Part II: Topical Studies in Oceanography 93, 16--32.
\url{https://doi.org/10.1016/j.dsr2.2013.01.008}

\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
Sosik, H.M., Schulz, J., Romagnan, J.-B., Picheral, M., Pearlman, J.,
Ohman, M.D., Niehoff, B., Möller, K.O., Miloslavich, P., Lara-Lpez, A.,
Kudela, R., Lopes, R.M., Kiko, R., Karp-Boss, L., Jaffe, J.S., Iversen,
M.H., Irisson, J.-O., Fennel, K., Hauss, H., Guidi, L., Gorsky, G.,
Giering, S.L.C., Gaube, P., Gallager, S., Dubelaar, G., Cowen, R.K.,
Carlotti, F., Briseño-Avena, C., Berline, L., Benoit-Bird, K., Bax, N.,
Batten, S., Ayata, S.D., Artigas, L.F., Appeltans, W., 2019.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00196}{Globally
consistent quantitative observations of planktonic ecosystems}.
Frontiers in Marine Science 6.

\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
Marra, G., Wood, S.N., 2011. Practical variable selection for
generalized additive models. Computational Statistics \& Data Analysis
55, 2372--2387. \url{https://doi.org/10.1016/j.csda.2011.02.004}

\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
Paired high-throughput, in situ imaging and high-throughput sequencing
illuminate acantharian abundance and vertical distribution. Limnology
and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}

\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
Johnson, R., Dickey, T.D., McNeil, J., Michaels, A.F., Knap, A.H., 1998.
Influence of mesoscale eddies on new production in the Sargasso Sea.
Nature 394, 263--266. \url{https://doi.org/10.1038/28367}

\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
and their symbionts. Marine Biology 97, 559--569.
\url{https://doi.org/10.1007/BF00391052}

\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
C.M., 1995. Planktonic sarcodines (Acantharia, Radiolaria, Foraminifera)
in surface waters near Bermuda: abundance, biomass and vertical flux.
Journal of Plankton Research 17, 131--163.
\url{https://doi.org/10.1093/plankt/17.1.131}

\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
Atlantic Time-series Study and the Hydrostation S program. Deep Sea
Research Part II: Topical Studies in Oceanography 43, 157--198.
\url{https://doi.org/10.1016/0967-0645(96)00004-5}

\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
K., Ogiso-Tanaka, E., 2023. DNA metabarcoding focused on
difficult-to-culture protists: An effective approach to clarify
biological interactions. Environmental Microbiology 25, 3630--3638.
\url{https://doi.org/10.1111/1462-2920.16524}

\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
Nakamura, Y., Suzuki, N., 2015.
\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_9}{Phaeodaria:
Diverse marine cercozoans of world-wide distribution}. Springer, pp.
223--249.

\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
resolved from planktonic organisms in situ. ICES Journal of Marine
Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}

\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
L., Hauss, H., Karp-Boss, L., Kiko, R., Lombard, F., McDonnell, A.M.P.,
Picheral, M., Rogge, A., Waite, A.M., Stemmann, L., Irisson, J.-O.,
2023. Three major mesoplanktonic communities resolved by in situ imaging
in the upper 500 m of the global ocean. Global Ecology and Biogeography
32, 1991--2005. \url{https://doi.org/10.1111/geb.13741}

\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
G., 2010. The Underwater Vision Profiler 5: An advanced instrument for
high spatial resolution studies of particle size spectra and
zooplankton. Limnology and Oceanography: Methods 8, 462--473.
\url{https://doi.org/10.4319/lom.2010.8.462}

\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
Setou, T., 2022. DNA metabarcoding reveals vertical variation and hidden
diversity of alveolata and rhizaria communities in the western north
pacific. Deep Sea Research Part I: Oceanographic Research Papers 183,
103765. \url{https://doi.org/10.1016/j.dsr.2022.103765}

\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
in the twilight zone: Their role in the carbon cycle. Limnology and
Oceanography 63, 2579--2594. \url{https://doi.org/10.1002/lno.10961}

\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00397}{The
roles of suspension-feeding and flux-feeding zooplankton as gatekeepers
of particle flux into the mesopelagic ocean in the northeast pacific}.
Frontiers in Marine Science 6.

\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
Suzuki, N., Not, F., 2015.
\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_8}{Biology
and ecology of radiolaria}. Springer, pp. 179--222.

\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
giant colonial radiolarian from equatorial and Gulf Stream waters. Deep
Sea Research Part A. Oceanographic Research Papers 28, 1033--1047.
\url{https://doi.org/10.1016/0198-0149(81)90016-9}

\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
role in silica transport to the deep sea. Science 222, 616--618.
\url{https://doi.org/10.1126/science.222.4624.616}

\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
zooplankton with the fine-scale vertical prey field. Limnology and
Oceanography 66, 3811--3827. \url{https://doi.org/10.1002/lno.11920}

\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
Wood, S.N., 2017. Generalized additive models: an introduction with R,
Second edition. ed, Texts in statistical science. CRC Press/Taylor \&
Francis Group, Boca Raton London New York.

\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.

\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians
and Their Role in Epipelagic Communities of the East Pacific Rise and
the Gulf of California. Biology Bulletin 32, 279--287.
\url{https://doi.org/10.1007/s10525-005-0103-5}

\end{CSLReferences}

\end{document}