### 2024-06-26 manuscript/01_intro.qmd
edit manuscript

---
bibliography: references.bib
---

# Introduction

Rhizaria are an extremely diverse super-group of single-celled organisms which occupy a wide range of habitats. Planktonic Rhizaria include Foraminifera, Radiolaria (including Acantharea and Collodaria) and Phaeodaria, all of which are observed throughout the global ocean [@biard2016; @laget2024] . While the taxonomy of these organisms has recently undergone several reclassifications [@biard2022], their presence in ocean ecosystems has been long known to oceanographers. Some of the earliest records of their existence are from oceanographic expeditions in the 19th century [@haekel1887]. Rhizaria are unique members of the plankton and protist community because they can reach large sizes (up to several mm in diameter) and they construct intricate mineral skeletons out of either silica, strontium, or calcium carbonate [@kimoto2015; @suzuki2015; @nakamura2015; @biard2022]. Despite their noticeable morphology and global distribution, Rhizaria were largely understudied throughout the 20th century. Fragile organisms like Rhizaria are difficult to adequately study because they can be destroyed through standard zooplankton sampling techniques and preserve poorly. A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @bijma1990; @caron1995]. However, the bulk of Rhizaria research was constrained to sediment traps or paleontological studies of sediment [@takahashi1983; @boltovskoy1993].

Recently, the advent of molecular techniques and *in situ* imaging tools have ignited a renewed focus on Rhizaria in pelagic ecosystems [@caron2016]. Molecular tools have shed light on Rhizaria taxonomy [@biard2022], distribution [@blanco-bercial2020; @sogawa2022], ecology [@decelle2012; @nakamura2023], biogeochemical roles [@gutierrez-rodriguez2019; @laget2024], and metabolism [@cohen2023]. Yet, despite the excellent taxonomic resolution provided by molecular approaches, they do not provide a truly quantitative metric for estimating Rhizaria abundance or biomass. *In situ* imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth2024]. @biard2016 utilized *in situ* imaging at a global scale to suggest large (\>500$\mu$m diameter) Rhizaria were substantial contributors to the ocean carbon standing stock. While more recent calculations suggest lower carbon contribution [@laget2024], Rhizaria nonetheless have substantial influences on biogeochemical cycling. This influence is due to their unique mineral skeletons and ability to repackage small particles into larger, fast sinking ones [@ikenoue2019]. Thus, some Rhizaria may play substantial roles in ocean export through their interception of sinking particles [@stukel2018; @laget2024]. However, Rhizaria ecology is poorly understood [@biard2022] and their ecological roles are likely more than diverse than simple particle interception.

The ecological role of Rhizaria in plankton communities is complicated due to the fact different taxa can exhibit very different trophic modes. Many Rhizaria are strictly heterotrophic [@biard2022], yet their feeding modes can be quite varied. Phaeodaria are largely thought to be flux-feeders, collecting and feeding on sinking particles [@gowing1989]. Alternatively, Retaria can be either exclusively heterotrophic or mixotrophic, utilizing photosynthetic algal symbionts [@anderson2014; @decelle2015]. Mixotrophic Foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and significantly contribute to total primary productivity [@kimoto2015]. Still, Foraminifera are omnivorous, possibly even predominately carnivorous, with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984]. Radiolaria have several lineages, many of which host symbionts [@biard2022b]. Amongst Radiolaria, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981]. All known Collodaria species host dinoflagellate symbionts [@biard2022b] and can contribute substantially to primary productivity, particularly in oligotrophic ocean regions [@caron1995; @dennett2002]. This Collodaria-symbiont association has been suggested as a reason for their high abundances throughout the photic zone of oligotrophic environments globally [@biard2016; @biard2017]. A few Acantharea (Radiolaria order) clades host algal symbionts [@decelle2012; @biard2022b], notably with two clades forming an exclusive relationship with *Phaeocystis*. However, globally, Acantharea are less abundant than Collodaria [@biard2022] and contribute less to total primary productivity [@michaels1995]. This may be due to the fact several clades of Acantharea are cyst-forming and strictly heterotrophs [@decelle2013; @biard2022b]. Furthermore, @marsbrisbin2020 documented apparent predation behavior in Acantharea near the surface, suggesting that there may be a larger reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among Rhizaria taxa, it is reasonable to expect some form of niche partitioning. A number of studies do suggest evidence for vertical zonation between Rhizaria groups according to various trophic strategies. Taxon-specific studies of Radiolaria suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017]. Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022], Phaeodaria, alternatively, are generally found beyond depths where light can sustain photosynthesis, but they can feed on sinking particles [@stukel2018; @laget2024]. In an imaging-based study of the whole Rhizaria community, @biard2020 noted clear patterns in vertical zonation which largely corresponded to different trophic roles. Protists communities, including Rhizaria, have been described to partition along autotroph-mixotroph-heterotroph gradients with increasing depth in the water column [@blanco-bercial2022; @laget2024]. Yet, few studies have made direct attempts to relate Rhizaria abundances to environmental factors [@biard2020]. In part, this is due to the fact few studies have been able to sample Rhizaria in the same location over a consistent timeframe [@michaels1988; @boltovskoy1993; @michaels1995; @hull2011; @gutiérrez-rodríguez2022]. Furthermore, very few studies have utilized *in situ* imaging, arguably the best method for quantifying Rhizaria, to consistently throughout the full mesopelagic [@laget2024] . Given this lack of information, there are many unknowns with respect to Rhizaria ecology, seasonality and phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria abundance measured for over a year from regularly occurring cruises at monthly intervals. We utilized an *in situ* imaging approach to facilitate abundance calculations. With this dataset, we address two critical aims. 1) Quantifying of large Rhizaria throughout the epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an annual cycle. These data were collected in the Sargasso Sea, and represents the first study of its kind in an oligotrophic system; and 2) We aim to test the hypothesis that Rhizaria exhibit niche partitioning according to trophic roles. This hypothesis has several implications, including vertical zonation, as seen in prior studies, but also that environmental variables related to trophic strategy will explain abundance patterns. Specifically, autotrophic/mixotrophic taxa will correspond to variables related to autotrophy (chl-a concentration, primary productivity, local DO maxima) and other rhizaria will correspond to factors which promote heterotrophy (particle concentration, flux, and local DO minima).