### 2024-06-26 manuscript/02_methods.qmd

---
bibliography: references.bib
---

# Methods

## Oceanographic sampling

Data were collected in collaboration with the Bermuda Atlantic Time-series Study [@michaels1996; @lomas2013] on board the R/V Atlantic Explorer. Cruises were conducted at approximately monthly intervals. Rhizaria individuals were sampled using the Underwater Vision Profiler 5 (UVP) [@picheral2010]. The UVP5 is an in-situ camera used to capture plankton and particle images and has been well established to accurately quantify large Rhizaria abundance [@biard2016; @stukel2018; @stukel2019; @biard2020; @barth2022; @drago2022; @llopismonferrer2022; @panaïotis2023]. The UVP5 was mounted to the sampling rosette and collected data autonomously on routine casts, from which only the downcast data are utilized. The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022, during which time the BATS region was sampled for 3-5 days at monthly intervals (see sampling details in Supplemental Table 1). Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda in the Sargasso Sea (approximately 31.0$^oN$-32.5$^oN$, 64.25$^oW$-63$^oW$; Supplemental Figure 1). In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m). However, Rhizaria were only typically found in large abundances throughout the epipelagic and mesopelagic zones. As such, we limit this study to results from the upper 1000m of the water column.

A variety of biotic and abiotic data were collected during each BATS cruise. The UVP5 provided particle count data at a high-frequency from each cast. Particle concentration, as a proxy for prey field [@whitmore2021], was calculated from this data for all particles 184$\mu m$ - 900$\mu m$. The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px; 0.092mm per pixel) and the upper size range was set to below the size of observable Rhizaria yet large enough to be inclusive of particles on which some Rhizaria may feed. For each UVP cast supporting continuous profiles of the CTD parameters salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen (DO), *in situ* chlorophyll fluorescence were measured at 24Hz using the BATS CTD package. On select casts, Niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths. On each cruise, flux estimates of total mass, organic carbon, and nitrogen were also collected using sediment traps; in the present study we utilized flux to the mesopelagic as the flux at 200m. Also primary productivity was estimated through measuring $^{14}C$ uptake rates from *in situ* incubations. For full descriptions of the BATS sampling program and methods, see @knap1997 and @lomas2013 for a review. Additionally, data can be viewed online [(https://bats.bios.asu.edu/bats-data/)](https://bats.bios.asu.edu/bats-data/).

Environmental data were processed to match the format of the Rhizaria abundance estimates (see below). CTD data were collected at higher frequency than the UVP (24Hz vs 15Hz respectively), so these data were averaged within matching bins to the UVP5 data. Data from Niskin bottles were first linearly interpolated in depth at 1m resolution then time averaged over the cruise, then subsequently averaged into matching bins as the UVP5 data. Primary productivity estimates were depth integrated throughout the euphotic zone (0-140m).

## Rhizaria imaging processing and abundance quantification

Individual vignettes of Rhizaria images were identified using the classification platform EcoTaxa [@picheral]. Data were pre-sorted utilizing a random-forest classifier and pre-trained learning set. Taxonomic classification were done based on morphological parameters measured in Zooprocess [@gorsky2010], which includes parameters such as major axis, equivalent spherical diameter (ESD), and grey values. While there are sparse taxonomic guides for *in situ* images of Rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020]. Using the aforementioned sources and publicly available EcoTaxa projects, we constructed a guide accessible at: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>. Broadly, Rhizaria were classified as Foraminifera, Radiolaria (Acantharea or Collodaria), or as a variety of Phaeodaria families (Figure 1). When identification could not be confidently made between a few candidate taxa, a less specific label was used. As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly Phaeodaria but not distinguishable into a family.

![Example images of different Rhizaria taxa as collected by UVP5 imaging. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)

The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$ (ESD). However, identified rhizaria ranged from a 934$\mu m$ ESD Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth2024]. However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded. The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals. Our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic (see @barth2022 for an extended discussion of UVP5 data processing). For the present study, Rhizaria abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic; Supplemental Table 1) while still maintaining ecologically relevant widths. However, concentrations in a 25m bin would need to be greater than 2.428 ind. $m^{-3}$ and 3.912 ind. $m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth2024]. Because we typically observed many Rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions. For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast. Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m). The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 ind. $m^{-2}$, 0.33 ind. $m^{-2}$, and 0.20 ind. $m^{-2}$ respectively. All UVP data processing was done using the `EcotaxaTools` package in R (Barth 2023).

## Modelling environmental controls of Rhizaria abundance

Generalized Additive Models (GAMs) were used to assess the relationship between integrated Rhizaria abundance and different environmental factors. GAMs offer the ability to model non-linear and non-monotonic relationships, which can be particularly useful in assessing ecological relationships [@wood2017] and have been successfully applied to Rhizaria ecology [@biard2020]. The `mgcv` package [@wood2001] was used to construct models relating environmental parameters to each taxonomic group's integrated abundance estimates from each cast. To select the most parsimonious model for each analysis, a backwards step-wise approach was taken. First, a full model was fit using any term which may be ecologically relevant. Terms were fit using maximum likelihood with a double penalty approach on unnecessary smooths [@marra2011]. The smoothness parameter was restricted (k = 6) to prevent over fitting the models. Then, the full model was checked for concurvity (a metric indicating high parameter relatedness in GAMs). If two terms had a high concurvity value (\>0.8), then two alternative models were constructed each leaving out one of the related terms. The more parsimonious model of these two was selected using AIC. Once concurvity of terms was reduced, then the model was subjected to the backwards step-wise procedure. At each iteration, the model term with the lowest F score (least statistically significant) was removed. This was repeated until all model terms were statistically significant or the $R^2_{adj}$ was substantially reduced. Models were fit for each water column region; epipelagic, upper mesopelagic, and lower mesopelagic. In cases where observations were too sparse for a given taxonomic grouping, models were not run. All code, full models, and model selection tools are available in open-source scripts, as well as intermediate data products at <https://github.com/TheAlexBarth/RhizariaSeasonality>.