### 2024-06-26 manuscript/03_results.qmd

---
bibliography: references.bib
---

# Results

### Environmental Variability

While the environmental conditions were generally low in variation and oligotrophic, there is considerable influence from deep winter-mixing and summer stratification as well as secondary influences from mesoscale eddies which result in spatiotemporal heterogeniety [@mcgillicuddy1998; @lomas2013]. Variability in the water column structure was visible during the study period (Figure 2). This is best evidenced through the temperature profiles; In the late summer and early fall there was a stratified water column with high temperatures in the surface (\<75m) (Figure 2A) and slightly elevated salinity (Figure 2B). This warm, stratified period appeared more intense during the few months sampled in 2019. In 2021, we observed the stratified layer slowly dissipated into the winter months following mixed layer entrainment. There was a consistent oxygen minimum zone (OMZ) located at about 800m deep (Figure 2C). February 2021 saw a notable downwelling event, likely due to a passing anti-cyclonic eddy which impacted the local region during early 2021. During this phase, warmer, oxic water was significantly depressed deeper into the mesopelagic. This process was reversed in the spring months (March, April) primary due to the interaction of convective mixing and the passing of a strong cyclonic eddy resulting in a deep cold mixed layer. Primary production was highest during the spring mixing period, evidenced both by *in situ* fluorescence (Figure 2D) and productivity incubation experiments (Figure 3A). Originating near the surface, the productivity peak moved deeper throughout the spring and declined into the summer (Figure 2D). However, there was a notable, yet smaller productivity bump in the late summer and early fall (Figure 3A) which occurred deeper in the epipelagic (Figure 2D). The particle concentration (184$\mu m$ - 900$\mu m$) was closely coupled to chlorophyll-a patterns.

![Environmental profiles across time-series of study period. Y axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved Oxygen. (D) in situ chlorophyll fluorescence. (E) Particle concenctrion (184 - 900\$\\mu m\$). (F) Bacteria Abundance. (G) Silica. (H) Nitrate](images/02_environmental.pdf)

Overall, particle concentration was high near the surface during the 2021 spring bloom, then moved deeper throughout the water column attenuating throughout the lower epipelagic (Figure 2E). Similarly, heterotrophic bacteria abundance was closely linked to overall productivity, althrough there was a more consistent moderate-abundance layer near the top of the mesopelagic (\~250m) (Figure 2F). Concurrent with the secondary fall production peak, there was also higher particle concentration and bacterial abundance in the later summer and early fall. Interestingly, while primary productivity estimates from July-August were not that different between 2019 and 2021 (Figure 3A), chlorophyll-a florescence, particle concentration, and bacterial abundance were much higher in 2019's summer/fall (Figure 2D-F). Inorganic nutrients ($Si$ and $NO_3$) were generally well stratified, with low concentrations in the epipelagic and increasing throughout the mesopelagic. However, both nutrients did vary vertically in accordance with the 2021 February downwelling and spring mixing period (Figure 2G-H). Additionally, in the late fall of 2021, $Si$ concentrations were slightly elevated in the mid-mesopelagic (Figure 2G).

![Primary productivity (A) and flux estimates from total mass (B), carbon (C), and nitrogen (D). Values are from monthly cruises with month displayed in corresponding colors. Absent data are shown by grey bar.](images/03_prod-flux.pdf)

Overall mass flux to the mesopelagic was highest during the 2021 February downwelling (Figure 3B). Generally, export was similarly high during March, declining in April then increasing slightly throughout the summer and early fall. While magnitude was slightly different, this pattern was consistent with total mass, carbon and nitrogen fluxes (Figure 3B-D). Higher mass, carbon and nitrogen flux also occurred in the 2019 late summer - early fall period.

### Rhizaria abundance and distribution

Across all imaged mesozooplankton (\>900$\mu m$), Rhizaria comprised a considerable fraction of the total community. Considering the total abundances of the observational period, Rhizaria comprised on average, 42.6% of all mesozooplankton abundance (Supplemental Figure 3). Copepods were the second most abundant, comprising 35.5% and all other living mesozooplankton were 22%. The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton abundance. In the mesopelagic Rhizaria were a smaller fraction, at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).

![Average abundance of Rhizaria in 25m bins, across entire study period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C), Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae (G), Coelodendridae (H).](images/04_vertical.pdf)

Average abundance of all Rhizaria had a bimodal distribution with respect to depth. Total abundance was highest just below the surface (0-100m), with secondary, wider peak occurring in the mid mesopelagic (Figure 4A). Variation in depth binned abundance was large, likely due to seasonal variability but also increased from the detection-risk described in the methods. The vertical distribution pattern and abundance varied considerably across taxonomic groups. Radiolaria were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5A). Acantharea displayed a bimodal distribution accounting for a large portion of the total Rhizaria pattern (Figure 4B, Figure 5). The surface layers were largely comprised of Collodaria, whose colonies were abundant (Figure 4C). The most abundant Phaeodaria, Aulacanthidae, also had a bimodal pattern with the density was highest in the lower mesopelagic (Figure 4D). Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower (Figure 4E). Aulosphaeridae had low average density and was nearly homogeneously distributed throughout the water column, although slightly lower in the epipelagic (Figure 4F). Castanellidae were the only Phaeodaria who appeared to be effectively restricted to the photic zone (Figure 4G). Alternatively, Coelodendridae primarily occurred in the lower mesopelagic (Figure 4H). A few individuals from the families Tuscaroridae and Medusettidae were also observed in the mesopelagic, yet they were much rarer (data not shown).

![Seasonality of Rhizaria integrated abundance for the epipelagic (0-200m) (A), upper mesopelagic (200-500m) (B), lower mesopelagic (500-1000m) (C). Data shown are the cruise mean integrated abundance in the water column region colored for each taxon.](images/05_seasonality.pdf)

Between the monthly cruises, Rhizaria integrated abundance varied in the epipelagic. Highest average abundance occurred in June 2021 and was lowest during the winter months (Figure 5A). The 2019 later summer - fall period also had much higher integrated abundance than similar months in 2021. While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrea abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5A). The mesopelagic integrated abundance was much more consistent across monthly cruises, although average abundance was notably higher in 2019 (Figure 5B, 5C). The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodaria families. However, Acantharea and unidentified Rhizaria also were common members of the community (Figure 5B, C).

### Body size throughout the water column.

Very few taxa had consistent distributions throughout the water column. Only Acanthrea, Foraminifera, Aulacanthidae, and Aulosphaeridae were consistently abundant in the epipelagic and mesopelagic. To investigate if the populations or morphologies shifted throughout the water column, we compared the sizes (ESD) between mesopelagic and epipelagic groups for each taxa. All groups were significantly different on average (Wilcox Rank Sum p-value \<0.001). Acantharea were smaller, on average in the mesopelagic while all other taxa tended to be larger (Figure 6).

![Comparison of average sizes (ESD) amongst Rhizaria taxa which occurred throughout the water column.](images/09_size-comp.pdf)

### Environmental Drivers of Rhizaria Abundance

![Partial effect plots from Generalized Additive Models (GAMs) of all Rhizaria abundance. Models were constructed separately for depth integrated abundance through the Epipelagic (0-200m) (A-G), Upper Mesopelagic (200-500m) (H-J), and the Lower Mesopelagic (500-1000m) (K-L). Each panel represents the partial effect of a GAM model term, with smoothed relationship shown shaded with 95% confidence intervals. Individual points are depth integrated values for all Rhizaria from an individual cast, colored by month.](images/06_tot-partials.pdf)

For integrated abundance of all Rhizaria, the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1). In the epipelagic, there were several significant predictor variables (Figure 7A-7G). This included terms which are likely indicators of changing water masses; Salinity (7E) and Temperature (7G). The model also had parameters related to both heterotrophy (Particle concentration, 7A) and autotrophy (Primary production, 7B). However the relationship to particle concentration is weakly increasing trend while the primary productivity effect appears to be influenced by a cruise with a high outlier for increased primary productivity. The models for the mesopelagic regions were much more reduced and the only significant effects were particle-related terms (particle concentration and mass flux). In the upper mesopelagic (Figure 7H-7J), final model included N flux, particle concentration and a non-significant effect for Silica. The lower mesopelagic (Figure 7K, 7L) only had model effects kept of total mass flux and particle concentration. For both the upper and lower mesopelagic, particle concentration had a clear, strong positive relationship (Figure 7I, 7L).

| **Model**                          | Term                   | edf    | F     | p       |
|---------------|---------------|---------------|---------------|---------------|
| **All Rhizaria Epipelagic**        | Temperature            | 0.7950 | 0.492 | 0.0440  |
| $R^2_{adj}$= 0.48                  | Salinity               | 2.4117 | 5.113 | \<0.001 |
|                                    | Silica                 | 2.1511 | 8.601 | \<0.001 |
|                                    | Bacteria Concentration | 0.8870 | 1.544 | \<0.001 |
|                                    | Avg Mass Flux          | 1.4107 | 0.802 | 0.0211  |
|                                    | Primary Productivity   | 1.9280 | 2.736 | \<0.001 |
|                                    | Particle Concentration | 0.9151 | 2.136 | \<0.001 |
| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6702 | 0.404 | 0.0814  |
| $R^2_{adj}$= 0.40                  | Avg N Flux             | 2.9524 | 3.172 | \<0.001 |
|                                    | Particle Concentration | 3.5568 | 11.82 | \<0.001 |
| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 2.036  | 2.953 | \<0.001 |
| $R^2_{adj}$= 0.61                  | Particle Concentration | 3.819  | 27.43 | \<0.001 |

: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column. edf = estimated degrees of freedom, F = F-statistic, p = p-value. Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.

GAMs for individual taxa were much less consistent in their fits (Table 2). This is likely in part due to the high number of non-observations for certain taxa. Note that due to low abundances, GAMs were not constructed for Tuscaroridae or Medusettidae. Furthermore no, significant terms were found for a model with Aulosphaeridae in the epipelagic nor Foraminifera in the mesopelagic.

Epipelagic Acantharea were explained by several predictor variables and had a good fit ($R^2_{adj}$ = 0.539, Table 2, Supplemental FIgure 4). Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.543). There were several significant explanatory variables, although the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Supplemental Figure 4U). Epipelagic Aulacanthidae similarly had several predictor variables which were significant, including both water quality parameters and particle/flux predictors (Table 2). There was a weak fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Supplemental Figure 4). Castanellidae also had a weak model fit in the epipelagic ($R^2_{adj}$ = 0.228), although particle concentration was the strongest predictor term (Table 2, Supplemental Figure 4).

| Model                                | Term                   | edf   | F     | p       |
|---------------|---------------|---------------|---------------|---------------|
| **Acantharea Epipelagic**            | Temperature            | 0.868 | 0.380 | 0.0935  |
| $R^2_{adj}$=0.539                    | Salinity               | 2.349 | 4.646 | \<0.001 |
|                                      | O2                     | 2.501 | 5.493 | \<0.001 |
|                                      | Avg Mass Flux          | 4.815 | 22.59 | \<0.001 |
|                                      | Bacteria               | 1.608 | 1.360 | 0.00945 |
|                                      | Particle Concentration | 0.877 | 1.414 | 0.00449 |
| **Acantharea Upper Mesopelagic**     | Avg Mass Flux          | 1.866 | 1.220 | 0.0166  |
| $R^2_{adj}$=0.197                    | Avg N Flux             | 2.208 | 3.100 | \<0.001 |
|                                      | Primary Productivity   | 2.382 | 2.497 | \<0.001 |
|                                      | Particle Concentration | 0.724 | 0.521 | 0.0437  |
| **Acantharea Lower Mesopelagic**     | Temperature            | 1.958 | 2.986 | \<0.001 |
| $R^2_{adj}$=0.496                    | Avg N Flux             | 1.513 | 1.565 | 0.0037  |
|                                      | Primary Productivity   | 0.856 | 1.153 | 0.0045  |
|                                      | Particle Concentration | 1.661 | 8.667 | \<0.001 |
| **Aulacanthidae Epipelagic**         | Temperature            | 2.591 | 5.003 | \<0.001 |
| $R^2_{adj}$=0.394                    | Salinity               | 1.523 | 1.789 | 0.002   |
|                                      | RFU                    | 0.872 | 1.315 | 0.0042  |
|                                      | Bacteria               | 0.808 | 0.833 | 0.0160  |
|                                      | Primary Productivity   | 2.622 | 9.160 | \<0.001 |
|                                      | Particle Concentration | 1.937 | 2.751 | \<0.001 |
| **Aulacanthidae Upper Mesopelagic**  | Avg C Flux             | 1.675 | 1.408 | 0.0109  |
| $R^2_{adj}$=0.132                    | Particle Concentration | 2.308 | 2.476 | 0.0013  |
| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.476 | 1.418 | 0.0076  |
| $R^2_{adj}$=0.308                    | Particle Concentration | 2.457 | 7.756 | \<0.001 |
| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.778 | 12.76 | \<0.001 |
| $R^2_{adj}$=0.196                    |                        |       |       |         |
| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 2.012 | 6.306 | \<0.001 |
| $R^2_{adj}$=0.148                    |                        |       |       |         |
| **Castanellidae Epipelagic**         | Salinity               | 0.915 | 0.400 | 0.0935  |
| $R^2_{adj}$=0.228                    | RFU                    | 0.912 | 2.052 | \<0.001 |
|                                      | Si                     | 2.018 | 3.656 | \<0.001 |
|                                      | Bacteria               | 1.128 | 0.883 | 0.0170  |
|                                      | Particle Concentration | 2.278 | 4.028 | \<0.001 |
| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.734 | 0.551 | 0.0527  |
| $R^2_{adj}$=0.09                     | Particle Concentration | 0.965 | 5.102 | \<0.001 |
| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 3.051 | 6.768 | \<0.001 |
| $R^2_{adj}$=0.157                    |                        |       |       |         |
| **Collodaria Epipelagic**            | Salinity               | 0.926 | 2.267 | \<0.001 |
| $R^2_{adj}$=0.16                     | O2                     | 2.015 | 2.217 | \<0.001 |
|                                      | Bacteria               | 1.843 | 2.100 | 0.0024  |
| **Foraminifera Epipelagic**          | Temperature            | 4.420 | 7.624 | \<0.001 |
| $R^2_{adj}$=0.543                    | O2                     | 0.799 | 0.788 | 0.023   |
|                                      | Bacteria               | 0.000 | 0.000 | 0.523   |
|                                      | Avg N Flux             | 1.847 | 3.022 | \<0.001 |
|                                      | Primary Productivity   | 0.472 | 0.139 | 0.183   |
|                                      | Particle Concentration | 1.685 | 2.955 | \<0.001 |

: Taxa-specific generalized additive models for different regions of the water column.edf = estimated degrees of freedom, F = F-statistic, p = p-value. Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.

In the upper mesopelagic (200-500m), abundances were generally low (Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae, Aulosphaeridae, and Coelodendridae (Table 2). All these models had generally poor fits ($R^2_{adj}$ \< 0.20). Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Supplemental Figure 5). The lower mesopelagic also had generally poor GAM fits for taxon specific models ($R^2_{adj}$ \<0.31), with the exception of Acantharea ($R^2_{adj}$ = 0.496). Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration (Supplemental Figure 6). For all Phaeodaria with a significant model, particle concentration was a main predictor variable (Table 2, Supplemental Figure 6).